---
title: "Fortune Global 500 Top 10 vs Countries' GDP"
author: "Mahesh Poudyal"
date: "11/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comparing country GDP with world's largest companies' annual revenue for 2019
This document outlines the steps and functions to produce global maps highlighting countries with Annual GDPs lower than those of the companies in [Fortune Global 500 top 10 list](https://en.wikipedia.org/wiki/Fortune_Global_500#Fortune_Global_500_list_of_2019) for the year 2019.

I'm using publicly available datasets and global maps from R package `rnaturalearth` to produce these maps using `ggplot2` and `sf` packages in R version 4.0.2.

### Loading required packages
```{r load_packages}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(readxl) #to read excel files
```


### Data preparation
I've raw data in `data` folder.
```{r data_prep}
#read gdp data
gdp <- read_xlsx("data/imf_weo_2019.xlsx", col_names = T)

str(gdp)

#read fortune global top 10 companies' data
gf500_top10 <- read_xlsx("data/fg500_2019_top10.xlsx", col_names = T)

#get countries data from rnaturalearth package into a dataframe
world <- ne_countries(scale = "medium", returnclass = "sf")

#add gdp data to world df
wd_gdp <- left_join(world, gdp,
                    by = c("geounit" = "Country"))

str(wd_gdp)

#add logical columns using whether gdp for a country is lower than that of top 10 companies' revenues
companies <- c(gf500_top10$Company) #getting list of company names to use in functions below

for (i in 1:length(companies)) {
  wd_gdp[,companies[i]] = as.factor(wd_gdp$`2019 (US$ billions)` < gf500_top10$`Revenues ($billions)`[gf500_top10$Company==companies[i]])
}

str(wd_gdp)
```


### Creating highlighted maps
I'll use `plots` folder to save maps created below.
```{r plots}
# function to create plots/maps
wd_gdp_plot = function(i) {
  ggplot(wd_gdp) + 
  geom_sf(aes(fill=.data[[i]])) + 
  theme(legend.position = "none",
        plot.title = element_text(colour = "orange", size = 24, face = "bold", hjust = 0.5),
        plot.caption = element_text(colour = "grey", size = 9, face = "italic")) + 
  scale_fill_manual(values = c("grey", "orange")) + 
    labs(title = paste0("Countries with Annual GDP lower than annual revenue of"," ",i),
         caption = "Data Sources: \n(i) Fortune Global 500 2019 (https://bit.ly/gf500-2019)\n(ii) IMF World Economic Outlook 2019 (https://bit.ly/imf-weo-2019)")
}

#test function
wd_gdp_plot(companies[7])


#save all plots/maps as list item using for loop
plot_list = list()
for (i in 1:length(companies)) {
  p=wd_gdp_plot(companies[i])
  plot_list[[companies[i]]] = p
}

#save plots as png, saves separate file for each plot.
for (i in 1:length(companies)) {
  file_name = paste("plots/wd_gdp_plot_", companies[i], ".png", sep="")
  png(file_name, width = 1000, height = 800)
  print(plot_list[[companies[i]]])
  dev.off()
}

```